---
title: "2024 Themes"
output: 
  html_document:
    fig_caption: FALSE
    includes:
      after_body: ../footer.html
css: "../indexstyle.css"
date: "March 1, 2024<br>Last update: `r format(Sys.time(), '%d %B, %Y')`"
---


::: {.container class="fluid-row d-flex flex-wrap align-content-stretch"}

```{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('diarymenu.html')
```
```

::::{.col-md-12}
Wow, it's been a while since I updated huh? I think I pushed out an update before I got to write up this blog post while I was testing deploy to neocities. At least if everything was working correctly.

Before I get into that: special shoutout to [scumsuck's](https://scumsuck.com/resources/deployneocities/) deploy-to-neocities guide. I also do not read good! So this helped me to understand what the hell I was looking at when the other guides didn't really help much.

With that out of the way... Whew. I didn't post at all last year, did I? To catch whatever few readers are still following along:

* I failed my first driver's test! (Practicing some more until I can make another attempt)
* I finished my thesis! I am not sharing it here for privacy reasons but if people want to know about VTubers and ambivalent sexism, send me a message? lol
* I graduated! After like 5 years in college :/
* Dealing with life with the... *few* bits of sanity I still have left.

Part of the reason I was not updating at all last year was definitely the depression and just having way too many projects and IRL commitments, not to mention keeping up with news as things around the world just...keep getting worse... I'm setting out to make this year a *somewhat* decent one though, even if I'm getting a rather late start on that? Oops. Well, that's just how it is with me. I remember last year I was crying to a friend at a party about how I'm just, god awful at juggling multiple things at once, and she nodded solemnly and said "Yeah, I noticed."

---

ANYWAY! All that nonsense aside, I've been thinking about my "theme" for this year which is something I picked up from the podcast [Cortex](https://www.relay.fm/cortex). I don't make 'resolutions' but 'themes', which are more like guiding words for a year and not hard numerical goals.

I've kind of settled on the idea of "motion" as my theme, because I want to move forward in my life...part of that is using RStudio to make sites -- I actually practiced first on [glitch](https://glitch.me) with a static site that I'm using as a "professional" portfolio. Obviously not linking that for security reasons, but I used a lot of what I learned from there to make my new site format... Hopefully I can write a post on that eventually (and why you *probably* shouldn't brute force a data science IDE to make a website like this).

---

My oracle cards from [Pippa Best's sea soul journey deck](https://seasoulblessings.com/sea-soul-journeys-oracle-cards/) for this year are as follows:

* Past: **Heard** -- May your brave call be carried on the wind
* Present: **Beauty** -- May you perceive nature's artistry afresh
* Future: **Imagine** -- May you look beyond what can be seen.

I'm choosing to interpet these as "you need to speak up more and be unapologetic about who you are", "live in the moment and appreciate what you can do, see, and create right now," and "the future is going to look different than what you imagined, but that's okay, and you'll still manage".

I finished my degree in psychology but a lot of my current feelings are leaning towards informatics in the future if I do continue my education, and I hope I can bring a unique perspective there? I care a lot about user experience and also how people interact with the internet... Well, only time will tell I guess.

::::

:::

::::{.container}

:::::{.col-md-12}

```{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('commentbox.html')
```
```
:::::

::::
