---
title: "Ditch Google Docs and Get a Blog"
output: 
  html_document:
    includes:
      after_body: ../footer.html
css: "../indexstyle.css"
date: "Created April 22, 2022<br>Last update: `r format(Sys.time(), '%d %B, %Y')`"
---

::: {.container class="fluid-row d-flex flex-wrap align-content-stretch"}

```{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('diarymenu.html')
```
```

::::{.col-md-12}

Soooo writing (fan)fiction and working on some of my interactive fiction projects has been hard. So here I am with another brain dump-y post that I was finally spurred into writing because of a lovely essay I saw on Twitter about butch and gnc girls in Japanese idol games...which I lost track of.

## Ditch the Google doc

Now obviously I'm comfortable with long form writing, but I think people can tell that my "blogs" are somewhat marred by my time spent on microblogging websites like Twitter and Tumblr. Don't get me wrong, I think microblogging has its place, but one thing has been aggravating me about the modern internet...and it's the overreliance on Google docs.

A lot of people are writing meta, posting complex ideas, or game documentation, and throwing it onto a Google doc and just hoping it doesn't randomly disintegrate. Now...this is all very "old man yells at cloud" moment here, but I have a few problems with this trend:

1. **Google Drive is known for deleting things randomly.** It has happened to two different friends of mine. This can happen for a variety of reasons, but I think this is enough for me to say "stop putting your interesting thoughts on Google docs and sharing it that way."
2. **It's just a bad reading experience.** Maybe this is just bias, but I have never had a good time reading a Google doc on my phone or on my laptop. I don't like it, it doesn't feel like I'm reading something for fun, because I always default to looking for errors or things I could improve. This might just be because of my specific use cases for google docs. Services like [publish gdocs](https://gdocs.pub) somewhat remedy this issue, though it still doesn't show up nice on mobile and it's still a pain for writing that people want to share with others quickly and easily, especially things they've put effort into.
3. **You are not immune to randomly deleting your own files.** Human error is real! It's happened to me! This is why I make sure to archive all my notes on worldbuilding somewhere else. Or in 3 places!

All this said: I understand the reliance on gdocs. I use it all the time for drafting ideas and sharing lyric translations...though I started just storing everything in simplenote and writemonkey recently because of how slow gdocs can get (my tools are probably not the best alternative either). But you see... I want to read more of what people have to say! In longform! And not just in small doses and bites. I want to have the ability to revisit their ideas! Some people online are very good writers and very good at synthesizing their ideas and they don't even realize it. Writing is an art form too, even non-creative writing!

I just think using google docs instead of blogs also takes away the fun... I love being able to remember a *short and descriptive* url for a blog post I found interesting, and to be able to read the person's other writing extremely easily. I care too intensely about this, though, and I'm preaching to the choir a bit since I'm on neocities.

## How do we return to blogging?
Ever since I started writing up things for blogs and seeing them more as [brain dumps](https://btxx.org/posts/dump/) instead of "impressive writing projects", it really lifted the weight on me. ~~Now if only this could work for my fiction writing~~. I'm sure some people will say, though, how do I blog? Where do I start? How do I even set one up? Luckily for you... I am an information hoarder and more than happy to open the door to the world of blogging for the intimidated. I won't be going too in depth since I just want to focus on the few things I like about the sites... I recommend just clicking on what interests you and seeing how you feel about it. If this spurs you on to start a blog, yay! If it doesn't though, that's ok too. (But if you make one...please leave a comment with a link to it, I love exploring the internet!) 

The following recommendations are generally assuming that you're fine with using a little bit of Markdown and basic HTML markup.

### Blog hosts 
These sites are really specifically focused on *blogging* and little else. Support for images might be limited and they're very focused on text. 

- **[bearblog](https://bearblog.dev/)** - I like this one a lot. It's free to sign up, but you can get more features with $25/year! The fact that it has an element of centralization is nice too, I always spend a lot of time browsing the new entries... Not to mention that the themes are lovely and responsive. I like that the developer is very committed to making software that lasts, and it's extremely easy to register your bearblog contents somewhere else if you need to move.
- **[mataroa](https://mataroa.blog/)** - Minimalist blogging only platform that focuses basically only on text, and you have to prove that you're a human dedicated to improving the internet™ when you sign up for a blog. I love their mission, but there's less "discoverability" like listed and bearblog have just in case that's important to you. This one is $9/year if you want to get extra features.

### Free static hosts
If you want to get your hands dirty and do a little bit of HTML writing, use an SSG, or use a template, these sites offer some fun free stuff.

- "old web" hosts
    - Ichi.city
    - Neocities
    - Nekoweb
- I've bundled all of these together because they all fundamentally do the same thing. They just provide an interface for you to upload a bunch of static files and then host it out there on the internet. I think this is good for people that *really* want to have ~old web~ vibes and maybe find (or even start) a webring of bloggers. Good ways to get started include [zoner](https://ryantrawick.itch.io/zoner) which is essentially a static generator based off of [zonelets](https://zonelets.net/), which is what I actually first started blogging with. [Static Ultra](https://scott2.neocities.org/projects/static-ultra/getting-started) is also a really amazing static generator tool that I don't see enough love for, and it comes with theme switching and good documentation from the dev...so if you're too intimidated and just want to start with something simple before making a whole site, I really recommend these! I'm only really familiar with Neocities of the stuff listed here and Nekoweb is relatively newer, but the developer [dimden](https://dimden.dev) is very dedicated to making a fun web experience :).

- **[glitch](https://glitch.com/)** - this is technically more techy than just a static host since they're for web app development, but their static hosting is free *and* they have an in browser code editor, so...why not...? Incidentally, this is what I use to host my "professional" site. They also have a great [eleventy template](https://glitch.com/edit/#!/remix/glitch-hello-eleventy) which is an SSG that lets you focus less on writing HTML and more on writing content, all in your browser. You can really just experiment on this without an account and see how you like it, too.

### Paid

If you're *just* doing blogging as a hobby and it's not your job, I honestly don't really recommend getting a paid service. A lot of sites host static stuff for free or dirt cheap, but these options are interesting and I like the features they offer. My experience with these is pretty limited, too.

- **[blot](https://blot.im/)** - this turns a Dropbox folder or a Google drive folder into a website. Pretty cool! It has theme support too, but I don't know how customizable it is. This one is $5/month.
- **[write.as](https://write.as/)** - they used to have a free tier, but not anymore, and they run $6/month.  You can also opt to self-host by using [writefreely](https://writefreely.org/).
- **[ghost](https://ghost.org/)** - Very popular alternative CMS to WordPress. I hear customer service is better and they're generally just faster and they get a plus for being open-source... I don't know much else. This one is $9/month for the basic plan.

### honorable mentions

Miscellaneous ones that I didn't know where to put them... I like exploring these though, so I don't think people should discount them!

- **[Dreamwidth](http://dreamwidth.org/)** - I'm not sure where to slot this one, honestly. I haven't used it personally, but it has a very strong fannish community if you're into that. A lot of fanfiction writers use this site, and a good number of people use dreamwidth to blog or post reviews. A good number of neocities folks also use this as a blog instead of hosting it directly on their site, which, is one way to do it I guess.
- [tiddlyhost](https://tiddlyhost.com) - Okay, so, this is TECHNICALLY for wikis, but you can *totally* use tiddlywiki or [featherwiki](https://feather.wiki/) as a blog. A lot of people use these to host blogs and if you like wiki-like structures (like me) this is actually a really great way to start with minimal code! Just explore the [hub](https://tiddlyhost.com/hub) for ideas! Featherwiki also has a [gallery](https://feather.wiki/?page=gallery#extensions) for ideas.
- [midnight.pub](https://midnight.pub/) - This is more like a message board than a blog, but I think it's still interesting. I love the interactivity and site culture here, and it's fairly anonymous.
- **[Everything else on the write.as alternative page here](https://alternativeto.net/software/write-as/?p=9)** - There's a lot of fun stuff here. I think telescope.as and pika.page look cool, but they're both freemium. Some of the stuff here isn't blogging focused though, so tread with caution.

*Note:* I did not include WordPress because I find it to be a bit overkill and antiquated in some ways...There are better options out there these days. I also did not include Tumblr because I don't think it's suitable for everyone, but it's still fun depending on what you make of it. Web builders like wix and carrd's blogging features are also not here because it's out of the scope of this blog post, but these all have a lot of tutorials out there online.

## my stuff isn't suitable for a blog, what do?

Well, it's true not everything is suitable for a blog. If you're documenting number-heavy things like a video game, I can't say a blog is the way you want to go. In these situations I would heavily recommend a wiki by using mediawiki or dokuwiki... ...sadly, I don't know anything about how to set these up. I'm just a wiki content mod that has nothing to do with the backend. That's too scary. If you don't have the resources to host, reliable wikifarms like miraheze and telepedia exist anyway. Just, [don't use fandomwiki](https://www.youtube.com/watch?v=qcfuA_UAz3I). For everyone's sake.

## I don't have anything interesting to say, though

I don't either. I don't think people on Twitter and etc. are necessarily more interesting than everyone either. I think it's important to embrace the "boring" parts of ourselves too! Giving yourself a voice is good and it's nice to practice putting your own thoughts into words, even if they're just floating around on the internet...

## why don't I just use a blogging platform like tumblr?

I definitely listed a few SNS-ish sites here, but I think the problem with Tumblr specifically is the imagined pressure to perform for people. You get this problem on anything that gets too big, Twitter is another beast. Personally, I don't think it's 100% feasible to escape this feeling, but I feel *much less* of this pressure on personal blogs... Sites like bearblog and mataroa feel less like the "town square" vibes of social media and more like an apartment? I realize this is a very strange comparison, but the Internet is basically the modern third space at this point.

## you forgot something!! or you got something wrong!!
I don't care. This is my blog. Write your own better blogging resource!!

Anyway go forth and blog! Be free to share your thoughts in longform! Who cares about being comprehensible!

::::

:::

::::{.container}

:::::{.col-md-12}

```{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('commentbox.html')
```
```
:::::

::::
